Dependent Object Types

### 2012-08-31 doc/candidacy/report.tex
show the scala ex... though not that insightful to.
little rev in ml section

Their module system addresses some shortcomings of the SML module
system. Though the SML module system is higher-order, because SML
provides functors and substructures, it still treats modules as
second-class. For example, it isn't possible to return a structure
from an if statement. In fact, allowing this might require obscuring
the visibility of the resulting structure to ensure soundness. Another
limitation of the SML module system is that ``type sharing'' is
limited to equality between type names, not general type
expressions. In effect, this means that the transparency / opaqueness
of a signature cannot be fully fine-tuned. Their proposal introduces
translucent signatures, which allows any type to be partially or fully
determined by a type expression.

@@

\subsubsection*{Pseudo-SML Example and Scala Encoding}
Consider this example, in a ML-like language implementing the calculus
of the paper. Since modules are first-class, we can \code{flip} a coin
and return a module depending on the outcome:
\begin{lstlisting}
structure XInt = struct
  type T = int
  val v = 3
  val f = negate
end
structure XBool = struct
  type T = bool
  val v = true
  val f = not
end
structure X =
  if flip() then XInt else XBool
\end{lstlisting}
But now, what should be the static type or signature of structure
\code{X}? In ML, in order to make the \code{if} type check, the two
branches must have equal types. In the calculus, the subsumption rule
is used to coerce both types to their least upper bound, represented
by the signature:
\begin{lstlisting}
signature XType = sig
  type T
  val v = T
  val f = T -> T
end
\end{lstlisting}

It is straightforward to encode this example in Scala, with the
correspondence Object = Module, and Object Type = Signature. Indeed,
the ${\nu}Obj$ calculus, which we study in section~\ref{sec:nuObj}
embraces this correspondence, and futher claims: Class = Method =
Functor, as we will see.

>>>>

Their module system addresses some shortcomings of the SML module
system. Though the SML module system is higher-order, because SML
provides functors and substructures, it still treats modules as
second-class. For example, it isn't possible to return a structure
from an if statement. In fact, allowing this might require obscuring
the visibility of the resulting structure to ensure soundness, which
we illustrate with an example below. Another limitation of the SML
module system is that ``type sharing'' is limited to equality between
type names, not general type expressions. In effect, this means that
the transparency / opaqueness of a signature cannot be fully
fine-tuned. Their proposal introduces translucent signatures, which
allows any type to be partially or fully determined by a type
expression.

@@

\subsubsection*{ML-like Example and Scala Encoding}
Consider this example, in a ML-like language implementing the calculus
of the paper. Since modules are first-class, we can \code{flip} a coin
and return a module depending on the outcome:
\begin{lstlisting}
structure XInt = struct
  type T = int
  val v = 3
  val f = negate
end
structure XBool = struct
  type T = bool
  val v = true
  val f = not
end
structure X =
  if flip() then XInt else XBool
\end{lstlisting}
But now, what should be the static type or signature of structure
\code{X}? In ML, in order to make the \code{if} type check, the two
branches must have equal types. In the calculus, the subsumption rule
is used to coerce both types to their least upper bound, represented
by the signature:
\begin{lstlisting}
signature XType = sig
  type T
  val v = T
  val f = T -> T
end
\end{lstlisting}

It is straightforward to encode this example in Scala, with the
correspondence Object = Module, and Object Type = Signature. Indeed,
the ${\nu}Obj$ calculus, which we study in section~\ref{sec:nuObj}
embraces this correspondence, and futher claims: Class = Method =
Functor, as we will see. Note that type members, and hence
path-dependent types, are crucial for this correspondence.
\lstinputlisting{ml.scala}
